---
title: "TMS_46v"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
The task involves two separate sessions: motor learning and working memory, counterbalanced.

(1) In the motor learning task, subjects randomly explored the space along an invisible arc in the left workspace. On the 16th trial, unbeknownst to them, we set the target by providing a boom. Subsequently, they have to keep repeating the movements to obtain more booms.

(2) In the working memory task, subjects arm were displaced in 3 different directions. They had to guess whether the test direction, following it, was one of the previously given directions. Only 2 lags were tested. WM score was computed as Hit - FA. There were 3 blocks during PRE and POST tms respectively.

```{r include=FALSE, cache=FALSE}

######### This file directly reads textfile from the experiment ############
rm(list = ls())
options(width=200)

require('ggplot2')
require('plyr')
require('Rmisc')
require('R.matlab')
require("reshape")

getwd()
dirpath <- "/Users/ae2012/Documents/R/TMS_data/"   # home directory in linux is shown as ~
setwd(dirpath)

## Easier to select which test version to use. This also determines filename to save.
myTest_version_string = "tms_param"   
#source("test_param.R")
source(paste(myTest_version_string,".R",sep=""))


allsubjects <- NULL
anchors <- 1: nANCHOR

##############################

for (i in 1 : nsubj ) {

  subjtab <- NULL
  
  for (mytask in tasks) {
  #print(mytask)
  
     for (block.num in 1 : nBLOCK.WM) {
       
         subjname <- s.names[i]
         my.file <- paste(substr(mytask,0,3), "_", subjname, block.num,".txt", sep="")
         my.file <- paste(dirpath,subjname,"_data/",my.file, sep="")

         # [Updated Feb 2018] I add error handler if Visual condition files aren't found.
         if (file.exists(my.file)) {
            tab <- read.table(file=my.file, sep=" ")
            # Remove NA columns; leaving only 8 usable columns!
            tab <- tab[,colSums(is.na(tab))==0]
            # Remove DELAY column, I don't need that!
            if (ncol(tab) > 7) { tab <- tab[,1:7] }
            # Then, give column names
            names(tab) <- c('trial',c(sprintf("anchor%i",anchors)),'probe','answer','RT') 
            ## Remove "ANSWER:" and get the Yes/No response...
            tab$answer <- sapply(tab[, nANCHOR + 3],
                                 function(x) {
                                   as.numeric(strsplit(as.character(x),":")[[1]][2])   # This is a list operation
                                 })
         }
         
         else {
            print(paste(subjname,"- File doesn't exist! NA will be used."))
            tab <- tab[FALSE,]  # Well this won't work if no prior tab df exists!
            tab[1,] <- c(rep(NA,11))
            tab$answer <- NA
         }
         tab$block   <- block.num
         tab$subj.id <- subjname
         tab$group   <- s.group[i]   # Control versus Treatment???
         subjtab <- rbind(subjtab, tab);   # Nice function to append data frame
     }
    
  }
  # Without doing this, function in ddply cannot distinguish the same trial number but of multiple blocks. 
  #subjtab$trial <- subjtab$block*24 + subjtab$trial.in.block
  subjtab$trial = 1: length(subjtab$subj.id==subjname)
  allsubjects <- rbind(allsubjects, subjtab)
  
}



## Make a column that represents for a given trial, whether the probe was one of the anchors or not
get.lag <- function(row) {
  ## Takes as input a row, and outputs a number that tells us which anchor was equal to the probe,
  ## or, if the probe was not among the anchors, then it returns NA.
  #index <- which(row$probe == row[,c('anchor1','anchor2','anchor3','anchor4','anchor5','anchor6')])
  index <- which(row$probe == row[,c(sprintf("anchor%i",anchors))])
  if (length(index)==0) { return (NA) } else { return (index-nANCHOR-1) }
  ## Ananda update: chronological order uses negative value, lag -6 means 6 directions ago from probe.
}



allsubjects$lag <- NA
## Go and check each row of our allsubjects data.
for (i in 1:nrow(allsubjects)) {
  allsubjects[i,]$lag <- get.lag(allsubjects[i,]);   ## Call the 'get.lag' function
}



# Creating 2 additional columns for correct answer and subject score!
allsubjects$lag <- as.factor(allsubjects$lag)
allsubjects$correct.answer <- ifelse( is.na(allsubjects$lag), 0, 1)
allsubjects$score <- ifelse( allsubjects$answer==allsubjects$correct.answer, 1, 0)


## For each of the anchors, compute the distance to the probe.
## NOTE: we invert the order of the anchors here. So dist1 refers to the distance
## to the probe of the *most recent* anchor (i.e. anchor 6), not the first anchor.
for (i in 1:nANCHOR) {
  allsubjects[,sprintf("dist%i",i)] <- abs(allsubjects$probe - allsubjects[,sprintf("anchor%i",nANCHOR+1-i)])
}

#str(allsubjects)



# TODO Ananda exercise: decide for each trial whether it is probe.outside or probe.inside,
# and then check whether proportion yes is different.
get.outside <- function(row) {
   ## Takes as input a row, and outputs a number that tells us when probe larger than maximum/minimum anchor
   #index <- which(row$probe > row[,c('anchor1','anchor2','anchor3','anchor4')])
   indexmax <- which(row$probe > max(row[,c(sprintf("anchor%i",anchors))]))
   indexmin <- which(row$probe < min(row[,c(sprintf("anchor%i",anchors))]))
   if (length(indexmax)>0 | (length(indexmin)>0)) { return (1) } else { return (NA) }
}

allsubjects$outside <- NA

for (i in 1:nrow(allsubjects)) {
   allsubjects[i,]$outside <- get.outside(allsubjects[i,])
}


###################

prop.yes <- ddply ( allsubjects, 
                    c("subj.id", "block", "lag", "group"),  # Note the grouping variables
                    function(xx) {
                        if (length(xx$answer)>1 && !is.na(xx$answer)) {
                           prop = sum(xx$answer)/length(xx$answer)
                           n = nrow(xx)
                           if ((sum(xx$answer) == length(xx$answer)) | (sum(xx$answer) == 0)) {
                              corrected = (sum(xx$answer) + 0.5)/(length(xx$answer)+1)
                           } 
                           else { corrected = prop }
                        }
                        else { 
                           prop = NA;  corrected = NA; n = 1 } # revised Feb 2018, for visual task with NA
                        return (c("uncorr" = prop,
                                  "prop.yes"= corrected,    ##### Corrected or uncorrected ??????
                                  "n" = n))
                    })


# NOTE: prop.yes contains both proportion of YES responses in trials when probe = anchor and probe = lures.
# We only want the Hit rate, that is, when probe = anchor. Next, we have to select just the lure trials 
# which we need to extract false alarm to compute hit minus false alarm.
#
no.lag <- subset(prop.yes, is.na(lag))

false.alarms <- data.frame(subj.id     = no.lag$subj.id,
                           block       = no.lag$block,
                           false.alarm = no.lag$prop.yes,    # Here, FA is based on corrected values
                           true.neg    = 1-no.lag$prop.yes)  # Specificity, true negative rate

# Merge two data frames!
#hit.fa <- merge(subset(prop.yes,!is.na(lag)), false.alarms)
hit.fa <- merge(subset(prop.yes,n!=12), false.alarms)


## Now compute WM performance;  (Revised Feb 2018 for visual WM with no data, NA).
## Which one is the best performance measure for WM?
hit.fa$hit.min.fa <- hit.fa$prop.yes- hit.fa$false.alarm                   # (1) Hit-FA rate, corrected
hit.fa$dprime <- qnorm(hit.fa$prop.yes) - qnorm(hit.fa$false.alarm)        # (2) dprime, corrected
hit.fa$beta <- exp(-qnorm(hit.fa$prop.yes)*qnorm(hit.fa$prop.yes)/2 +      # (3) beta, corrected
                    qnorm(hit.fa$false.alarm)*qnorm(hit.fa$false.alarm)/2)


```

#### Analyses on WM Scores Per Lag
```{r echo=FALSE, cache=FALSE}
pre.post <- ddply(hit.fa, c("subj.id","lag","group"), function(xx){
                  pre  <- subset(xx, block <=3 )
                  post <- subset(xx, block > 3 )
                  return(data.frame( "pre.wm"  = mean(pre$hit.min.fa, na.rm=TRUE),
                                     "post.wm" = mean(post$hit.min.fa, na.rm=TRUE),
                                     "pre.hit" = mean(pre$prop.yes, na.rm=TRUE),
                                     "post.hit"= mean(post$prop.yes, na.rm=TRUE),
                                     "pre.fa"  = mean(pre$false.alarm, na.rm=TRUE),
                                     "post.fa" = mean(post$false.alarm, na.rm=TRUE) ))
} )


##### Let's melt the data frame
pre.post.melt <- melt(pre.post, id.vars = c("subj.id","lag","group"), 
                      measure.vars = c("pre.wm", "post.wm","pre.hit", 
                                       "post.hit","pre.fa", "post.fa"))
colnames(pre.post.melt)[4] = "time"
colnames(pre.post.melt)[5] = "score"

# (1) Compute hit minus false alarm of GROUP DATA!
grp.lag.hit.fa <- ddply(pre.post.melt, c("lag","group","time"), 
                function(xx){
                     mean = mean(xx$score)
                     stde = sd(xx$score)/sqrt(length(xx$score))
                     new.var = as.character(xx$time)
                     time = unlist(strsplit(new.var, split="\\."))[1]
                     var  = unlist(strsplit(new.var, split="\\."))[2]
                     return(data.frame("time"=time,"n"=nrow(xx),
                                       "param"=var,"mean"=mean,"se"=stde))
})

# Plot the graph for each score......
ggplot( grp.lag.hit.fa, aes(x=time, y=mean,
                        ymax=mean+se, 
                        ymin=mean-se,
                        group=lag,
                        col=lag) ) + 
  geom_point() + geom_line() +
  geom_errorbar(width=0) +
  facet_grid(param~group,scales="free_y") +
  theme_bw()
print("Working memory score is denoted as wm = hit - fa")

```

#### What if I collapse the lags by taking the average between the two lags?!

```{r echo=FALSE, cache=FALSE}

# (2) What if we disregard the lags!?
grp.hit.fa <- ddply(pre.post.melt, c("group","time"), 
                    function(xx){
                      mean = mean(xx$score)
                      stde = sd(xx$score)/sqrt(length(xx$score))
                      new.var = as.character(xx$time)
                      time = unlist(strsplit(new.var, split="\\."))[1]
                      var  = unlist(strsplit(new.var, split="\\."))[2]
                      return(data.frame("time"=time,"n"=nrow(xx),
                                        "param"=var,"mean"=mean,"se"=stde))
                    })

ggplot( grp.hit.fa, aes(x=time, y=mean,
                        ymax=mean+se, 
                        ymin=mean-se, group = group) ) + 
  geom_point() + geom_line() +
  geom_errorbar(width=0) +
  facet_grid(param~group,scales="free_y") +
  theme_bw()
```


```{r echo=FALSE, warning=FALSE, cache=FALSE}
# (3) Perform ANOVA......
require("ez")

ezANOVA( data=subset(pre.post.melt, pre.post.melt$time %in% c("pre.wm","post.wm")),
         dv=score,
         wid=subj.id,
         within=time,
         between=group )
         
```

#### What if instead of doing ANOVA I do one-tailed t-test?
For each subject, I computed the difference between POST minus PRE in both groups. Then, the mean of this difference will be compared, between the treatment and sham groups.

```{r echo=FALSE, cache=FALSE}

#(5) What about finding the difference?
pre.post$wm <- pre.post$post.wm-pre.post$pre.wm
pre.post$wm.norm <- pre.post$wm/pre.post$pre.wm

pre.post$hit <- pre.post$post.hit-pre.post$pre.hit
pre.post$hit.norm <- pre.post$hit/pre.post$pre.hit

pre.post$fa <- pre.post$post.fa-pre.post$pre.fa
pre.post$fa.norm <- pre.post$fa/pre.post$pre.fa

pre.post.diff <- pre.post[,-which(names(pre.post) %in% c( "pre.wm","post.wm",
                                                          "pre.hit","post.hit",
                                                          "pre.fa","post.fa"))]

diff.score <- melt(pre.post.diff, id.vars = c("subj.id","lag","group"),
                   measure.vars = c("wm.norm","wm",
                                    "hit.norm","hit",
                                    "fa.norm","fa"))
colnames(diff.score)[4] = "time"
colnames(diff.score)[5] = "score"


grp.diff <- ddply(diff.score, c("group","time"), 
                        function(xx){
                          mean = mean(xx$score)
                          stde = sd(xx$score)/sqrt(length(xx$score))
                          return(data.frame("n"=nrow(xx),"mean"=mean,"se"=stde))
                        })

# Plot the graph for each score. Position dodge is to make the bars side-by-side.
# Note: What is width argument in position_dodge()?
ggplot(  subset(grp.diff, grp.diff$time %in% c( "wm","hit","fa")), 
        aes(x=group,y=mean,ymax=mean+se,ymin=mean-se,fill=group )) + 
  geom_bar(stat="identity", position=position_dodge(width=0.9))  +
  geom_point(aes(y=mean), position=position_dodge(width=0.9)) +
  geom_errorbar(stat="identity", width=0,position=position_dodge(width=0.9)) +
  facet_grid(.~time,scales="free_y") +
  #ylim(-0.1,0.05)+
  theme_bw()

diff.score.avg <- ddply(diff.score, c("subj.id","group","time"), 
                                 function(xx){
                                   mean = mean(xx$score)
                                   stde = sd(xx$score)/sqrt(length(xx$score))
                                   return(data.frame("n"=nrow(xx),"mean"=mean,"se"=stde))
                                 }) 
diff.score.avg <- subset(diff.score.avg, diff.score.avg$time=="wm")

t <- t.test(subset(diff.score.avg, diff.score.avg$group=="tms")$mean, 
       subset(diff.score.avg, diff.score.avg$group=="sham")$mean,
       alternative = c("less"))
sprintf("Difference in WM scores, Sham vs TMS group, p = %f",t$p.value)

```

### Kinematics Analyses of Motor Learning

```{r echo=FALSE, cache=FALSE, fig.width=6, fig.height=6}
##############################################
############### MOTOR LEARNING ###############
##############################################
getwd()
options(width=200)
library(corrplot)

dirpath <- "~/Documents/R/TMS_data/"   # home directory in linux is shown as ~
setwd(dirpath)

## Easier to select which test version to use. This also determines filename to save.
myTest_version_string = "tms_param"   
source(paste(myTest_version_string,".R",sep=""))

allsubjects <- NULL

for ( i in 1 : nsubj ) {
  subjtab <- NULL
  for ( block.num in 1 : nBLOCK.MOTOR ) {
     subjname <- s.names[i]
     group <- s.group[i]
     my.file <- paste("motorLog_", subjname, '_', block.num,".txt", sep="")
     my.file <- paste(dirpath,subjname,"_data/",my.file, sep="")
     # Load textfile as table, very easy in R. Note the column separator!
     tab <- read.table(file=my.file, sep=",", header=TRUE)
     tab$subj.id <- subjname
     tab$group   <- group
     tab$block   <- block.num
     tab$abs.angle <- abs(tab$angle_maxv_shift)
     tab$success <- ifelse(tab$boom,"success","failure")
     nrow(tab)
     # Nice function to append (COMBINE) several data frames together.
     subjtab <- rbind(subjtab, tab)
  }
  subjtab$newTrial = 1:nrow(subjtab)
  allsubjects <- rbind(allsubjects, subjtab)
}


# (1) Group average for the individual trial....
# Exclude the pre-train block...
my.df <- ddply( subset(allsubjects, task!="pre_train"), "subj.id", function(xx){
                      xx$newTrial = 1:nrow(xx)
                      xx$angle_maxv_shift = abs(xx$angle_maxv_shift)
                      return(xx) }
)

error.grp <- summarySE( my.df, "abs.angle", c("newTrial","task","group"), na.rm = TRUE)
error.grp <- subset(error.grp, error.grp$newTrial > 1)

g<-ggplot(error.grp, aes( x=newTrial,y=abs.angle,
                       ymin=abs.angle-se,
                       ymax=abs.angle+se,
                       color=task) ) +
  scale_color_manual(values=c("training"="blue","motor_post"="red")) +
  scale_fill_manual (values=c("training"="blue","motor_post"="red")) +
  geom_point() + geom_errorbar(width=0) +
  geom_smooth(method="lm") +
  #geom_ribbon( aes(fill=task), alpha=0.4 ) +
  ylim(0,20) + 
  ggtitle("Accuracy performance across all subjects") +
  xlab("Trial number") + ylab("Mean Absolute Deviation (degree)") +
  facet_grid(group~.) +
  theme_bw() + theme( 
    panel.background = element_rect(fill = "transparent") # bg of the panel
    , plot.background = element_rect(fill = "transparent") # bg of the plot
    , panel.grid.major = element_blank() # get rid of major grid
    , panel.grid.minor = element_blank() # get rid of minor grid
  )

ggsave( sprintf('David-grant.pdf'),g)


### Get learning slope? Using ddply with a predefined function.
get.learn.slope <- function(xx) {
  train.blocks <- subset(xx, task=="training")
  tot_trial    <- seq(1, length(train.blocks$angle_maxv_shift))
  yy <- lm(abs(train.blocks$angle_maxv_shift) ~ tot_trial)
  yy <- yy$coefficients[2]
  return (yy)
}

lin.slope <- ddply(allsubjects, c("subj.id","group"), get.learn.slope)

t <- t.test(subset(lin.slope,group=="sham" )$tot_trial,
       subset(lin.slope,group=="tms")$tot_trial,
       alternative = c("less"))

sprintf("Slope Difference of Linear Fit, Sham vs TMS group, p = %f",t$p.value)





# (2) Average accuracy motor post
post.grp <- subset(allsubjects, task=="motor_post")
#post.grp <- summarySE( post.grp, "abs.angle", c("task","group"), na.rm = TRUE)
t <- t.test(subset(post.grp,group=="tms" )$abs.angle,
       subset(post.grp,group=="sham")$abs.angle)

sprintf("Accuracy difference during Motor Post, Sham vs TMS group, p = %f",t$p.value)

```

The stats follow. The first part is the t-test comparing total reinforcement. The second part is comparing average absolute deviation during POST.

```{r echo=FALSE, cache=FALSE, fig.width=8, fig.height=4}
# (3) Total reinforcement.....
reward <- summarySE(my.df, "boom", c("subj.id","group"))
#reward <- summarySE(reward, "boom", c("group"))
t <- t.test(subset(reward,group=="tms" )$boom,
       subset(reward,group=="sham")$boom)

sprintf("Difference in total boom, Sham vs TMS group, p = %f",t$p.value)


# (4) Total reinforcement for different blocks.....
reward.blk <- summarySE(my.df, "boom", c("subj.id","group","block"))
reward.blk.avg <- summarySE(reward.blk, "boom", c("group","block"))
reward.blk.avg <- subset(reward.blk.avg, reward.blk.avg$block<5)

# Then draw the plot to compare two groups...
ggplot(subset(reward.blk.avg, block%in%c(1,4)),
       aes(x=block,y=boom,ymin=boom-se,ymax=boom+se,fill=group) ) +
  geom_bar(stat="identity", position=position_dodge())  +
  geom_point(aes(y=boom), position=position_dodge(width=2.9)) +
  geom_errorbar(stat="identity",width=0,position=position_dodge(width=2.9)) +
  xlab("Block") + ylab("Reward") +
  theme_bw()
#ggsave("plot.pdf",g)
```
What if you compute the difference in reward obtained between block 1 and 4. 
TMS group versus Sham group in terms of reward increment

```{r echo=FALSE, cache=FALSE, fig.width=8, fig.height=4}
### What if we compute the difference between block 1 and block 4.
reward.diff <- ddply(reward.blk, c("subj.id","group"), function(xx){
                  early = subset(xx, block==1)
                  late  = subset(xx, block==4)
                  c("diff"=late$boom - early$boom) }
                  )

t <- t.test(subset(reward.diff,group=="tms" )$diff,
       subset(reward.diff,group=="sham")$diff,
       alternative = c("less"))

sprintf("Reward increment between Sham vs TMS group, p = %f",t$p.value)


```

#### Change in movement direction contingent upon current trial outcome.....
Trial-to-trial change in movement direction was also compared between the TMS and sham group. The plot below includes 4 training blocks and motor post without reinforcement (so it is shown as "failure").

```{r echo=FALSE, cache=FALSE, fig.width=8, fig.height=4}
# (5) Average delta.m
d.m <- ddply( allsubjects,  c('subj.id','task',"group"),
              function(xx) {
                xx <- xx[ order(xx$newTrial), ]  # ensuring the order
                abs.dm<-NULL; delta.m<-NULL; next.boom<-NULL
                for (i in 1:(nrow(xx)-1)) {
                  # Accessing each trial; keep adding element to variable "delta.m"
                  delta.m  <- c(delta.m,    xx[i+1,"angle_maxv_shift"]-xx[i,"angle_maxv_shift"])
                  abs.dm   <- c(abs.dm, abs(xx[i+1,"angle_maxv_shift"]-xx[i,"angle_maxv_shift"]))
                  next.boom<- c(next.boom, xx[i+1,"boom"])
                }
                delta.m  <- c(delta.m, NA); abs.dm <- c(abs.dm, NA)
                return(data.frame(trial=xx$newTrial,
                                  block=xx$block,
                                  session=unique(xx$session),
                                  delta.m=delta.m,
                                  abs.dm = abs.dm,
                                  next.boom = c(next.boom,0),
                                  success=xx$success )) 
              }
)
avg.trial.trial <- ddply( d.m, c("task","success","group","block"), 
                          function(xx){
                            n.subj = length(unique(d.m$subj.id))
                            c( avg = mean(xx$abs.dm, na.rm = TRUE),
                               N  = length(xx$abs.dm),
                               n.subj  = n.subj,
                               se = sd(xx$abs.dm, na.rm = TRUE)/sqrt(n.subj)) # Note!
                          })

# Create clearer naming to indicate reward status for the training phase.
avg.trial.trial$task.block = ifelse (avg.trial.trial$task=="training",
                                      sprintf("%s_%s", avg.trial.trial$task, avg.trial.trial$block), 
                                      as.character(avg.trial.trial$task))
avg.trial.trial$task.block <- as.factor(avg.trial.trial$task.block)

# Reorder the levels according to the following task phase....
task = c("pre_train", "training_1","training_2","training_3","training_4","motor_post")
avg.trial.trial$task.block <- factor(avg.trial.trial$task.block, levels = task)

avg.trial.trial <- subset(avg.trial.trial, avg.trial.trial$task!="pre_train")

ggplot(avg.trial.trial, aes(x=task.block, 
                            y=avg, 
                            ymin=avg-se,
                            ymax=avg+se,
                            group=success,
                            color=success)) + 
  geom_point() + geom_line() +
  geom_errorbar(width=0) + 
  theme(plot.title = element_text(hjust = 0.5)) +
  ggtitle("Average magnitude of change in movement direction") +
  ylab("Angle (degree)") + xlab("") +
  facet_grid(.~group) +
  theme_bw() + theme( 
    panel.background = element_rect(fill = "transparent") # bg of the panel
    , plot.background = element_rect(fill = "transparent") # bg of the plot
    , panel.grid.major = element_blank() # get rid of major grid
    , panel.grid.minor = element_blank() # get rid of minor grid
    
  )
```




